---
title: "715_project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Set working directory (MAC)
# setwd("/Users/isabellachittumuri/Desktop/Project/STAT 715/PRSA_Data_20130301-20170228")

#Packages
library(dplyr)
library(tidyverse)
library(readxl)
library(readr)
library(tidyverse)
# library(qualityTools)
```

# Check which multi-site to use 

```{r eval=FALSE, include=FALSE}
#Import all cvs for 12 cites 
df1 <- read.csv("PRSA_Data_Aotizhongxin_20130301-20170228.csv", header = T)
df2 <- read.csv("PRSA_Data_Changping_20130301-20170228.csv", header = T)
df3 <- read.csv("PRSA_Data_Dingling_20130301-20170228.csv", header = T)
df4 <- read.csv("PRSA_Data_Dongsi_20130301-20170228.csv", header = T)
df5 <- read.csv("PRSA_Data_Guanyuan_20130301-20170228.csv", header = T)
df6 <- read.csv("PRSA_Data_Gucheng_20130301-20170228.csv", header = T)
df7 <- read.csv("PRSA_Data_Huairou_20130301-20170228.csv", header = T)
df8 <- read.csv("PRSA_Data_Nongzhanguan_20130301-20170228.csv", header = T)
df9 <- read.csv("PRSA_Data_Shunyi_20130301-20170228.csv", header = T)
df10 <- read.csv("PRSA_Data_Tiantan_20130301-20170228.csv", header = T)
df11 <- read.csv("PRSA_Data_Wanliu_20130301-20170228.csv", header = T)
df12 <- read.csv("PRSA_Data_Wanshouxigong_20130301-20170228.csv", header = T)

#Only keep columns: year, month, day, hour, PM2.5
df1 <- df1[c(2:6)]
df2 <- df2[c(2:6)]
df3 <- df3[c(2:6)]
df4 <- df4[c(2:6)]
df5 <- df5[c(2:6)]
df6 <- df6[c(2:6)]
df7 <- df7[c(2:6)]
df8 <- df8[c(2:6)]
df9 <- df9[c(2:6)]
df10 <- df10[c(2:6)]
df11 <- df11[c(2:6)]
df12 <- df12[c(2:6)]

# check for missing data
sum(is.na(df1))
sum(is.na(df2))
sum(is.na(df3))
sum(is.na(df4))
sum(is.na(df5))
sum(is.na(df6))
sum(is.na(df7))
sum(is.na(df8))
sum(is.na(df9))
sum(is.na(df10))
sum(is.na(df11))
sum(is.na(df12))
```

# Exploratory Analysis

Wanliu Cite 11 (PRSA_Data_Wanliu_20130301-20170228.csv) has the least amount of missing data with only 382 N.A. values

```{r}
wanliu <- read.csv("PRSA_Data_Wanliu_20130301-20170228.csv", header = T)
og_wanliu <- wanliu[c(2:6)]
```

```{r}
# combine year, month, day, hour
# in UTC (Coordinated Universal Time)
library(lubridate)
og_wanliu$Date <- with(og_wanliu, ymd_h(paste(year, month, day, hour, sep= ' '))) 
wanliu <- og_wanliu[c(5,6)] # only take pm2.5 and new date column
```

```{r}
# Precentage of NA values
nrow(wanliu) # 35064
summary(wanliu) # 382 NAs
# 382/35064 = 0.01089436 (about 1 precent)
```

The NAs only make up 1% of the data.

```{r}
# wanliu <- na.omit(wanliu) # omit missing values 
summary(wanliu)
str(wanliu)
class(wanliu)
```

Date column is in POSIXct format. 
We can see that there is a huge difference between median and maximum values.

# Check for outliers

## Before Aggregation

```{r}
library(gridExtra)

p1 <- wanliu %>%
  ggplot(aes(x = PM2.5)) +
  geom_boxplot(color = "brown", fill = "white") + labs(title="Distribution of PM2.5",x='ug/m3',y='PM2.5') 

p2 <- wanliu %>%
  ggplot(aes(x = PM2.5)) +
  geom_histogram(color = "brown", fill = "white", binwidth=20)

grid.arrange(p1,p2 ,nrow=2)
```

From above, we find that there are too many outliers found on these pollutants. We cant say them as outlier instead can call them as high unexpected value. 

## After Aggregation

Take the maximum value for each day and averaged it for each month

Why averaged max instead mean?
- to see maximum cause done by air pollutant per day.
- there can be peak hours in a day where values can go higher and drop down all of sudden. If we average for all those values from a day,it could be non relevant

```{r}
max_pollutants_per_day<-wanliu %>%
     as.data.frame %>%
    select('Date','PM2.5')%>%
      mutate(date_ymd=as.Date(Date,format="%Y-%m-%d"))%>%
           group_by(date_ymd)%>%
         summarise(PM2.5=max(PM2.5,na.rm=TRUE))

daily_max <- as.data.frame(max_pollutants_per_day)
```

```{r}
p1 <- daily_max %>%
  ggplot(aes(x = PM2.5)) +
  geom_boxplot(color = "goldenrod3", fill = "white") + labs(title="Distribution of PM2.5 on Daily Max Values",x='ug/m3',y='PM2.5') 

p2 <- daily_max %>%
  ggplot(aes(x = PM2.5)) +
  geom_histogram(color = "goldenrod3", fill = "white", binwidth=20) + labs(x='ug/m3',y='Count')

grid.arrange(p1,p2 ,nrow=2)
```

After aggregating, we could see that the plot are bit more normalized comparative to previous.But still we find some outliers in this case.

# Time series flow

- From previous section, we saw the distribution of four pollutants. Now we will see the time series flow for the same.

```{r}
library(scales)
# using hourly data for all years
ggplot(data = wanliu, aes(x = Date, y = PM2.5)) + 
  geom_line(color = "#00AFBB") + xlab('Date')+ylab('ug/m3')+ggtitle(paste('Hourly Emission of PM2.5'))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")

# using daily max
ggplot(data=daily_max,aes(x=date_ymd,y=PM2.5))+geom_line(color="lightslateblue")+
scale_x_date(breaks = seq(as.Date("2013-01-01"), as.Date("2017-01-01"), by="6 months"),labels = date_format("%b-%y"))+
xlab('Date')+ylab('ug/m3')+ggtitle(paste('Daily Maximum Emission of PM2.5'))+
theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")

# getting monthly avg
month_avg<-daily_max%>%
    mutate(y_m_d=paste(substr(date_ymd,1,nchar(date_ymd)+2),"-01"))%>%
    group_by(y_m_d)%>%
   summarise(avg_emission=mean(PM2.5,na.rm=TRUE))

ggplot(data=month_avg,aes(x=as.Date(y_m_d, format = "%Y - %m - %d"),y=avg_emission))+geom_line(color="magenta3")+
scale_x_date(breaks = seq(as.Date("2013-01-01"), as.Date("2017-01-01"), by="6 months"), date_labels = "%b-%y")+
xlab('Date')+ylab('ug/m3')+ggtitle(paste('Monthly Average Emission of PM2.5'))+
theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")
```

# Monthly Series for all years

- Taken the maximum value of each day and averaged it for all months.The values are scaled or performed log.We saw the same plots in previous section. Lets have a look at it all at a time.

```{r eval=FALSE, include=FALSE}
ggplot(data=month_avg,aes(x=as.Date(y_m_d,format="%Y - %m - %d"),y=scale(avg_emission)))+ geom_line(color="lightslateblue")+
    scale_x_date(breaks = seq(as.Date("2013-01-01"), as.Date("2017-01-01"), by="12 months"), date_labels = "%Y")+
xlab('')+ylab("PM2.5")+ ggtitle("Monthly Averaged on Daily Maximum Values") + 
theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")   
```

# Hourly Series

- Series contains maximum value recorded for each hour

```{r eval=FALSE, include=FALSE}
max_each_hour<-og_wanliu%>%
    select('hour','Date','PM2.5')%>%
  group_by(hour)%>%
  summarise(count=max(PM2.5,na.rm=TRUE))

p1<-ggplot(max_each_hour,aes(as.numeric(hour),count))+geom_line(color="magenta3")+
    scale_x_continuous(breaks=seq(0,23,1), limits=c(0,23)) +
    xlab('')+ylab('PM2.5') + ggtitle("Max for Each Hour")

mean_each_hour<-og_wanliu%>%
    select('hour','Date','PM2.5')%>%
  group_by(hour)%>%
  summarise(count=mean(PM2.5,na.rm=TRUE))

p2<-ggplot(mean_each_hour,aes(as.numeric(hour),count))+geom_line(color="magenta3")+
    scale_x_continuous(breaks=seq(0,23,1), limits=c(0,23)) +
    xlab('')+ylab('PM2.5') + ggtitle("Mean for Each Hour")

min_each_hour<-og_wanliu%>%
    select('hour','Date','PM2.5')%>%
  group_by(hour)%>%
  summarise(count=min(PM2.5,na.rm=TRUE))

p3<-ggplot(min_each_hour,aes(as.numeric(hour),count))+geom_line(color="magenta3")+
    scale_x_continuous(breaks=seq(0,23,1), limits=c(0,23)) +
    xlab('')+ylab('PM2.5') + ggtitle("Min for Each Hour")

grid.arrange(p1,p2,p3)
```

- shows there is not set time, that pollution occurs 

```{r eval=FALSE, include=FALSE}
max_each_month<-og_wanliu%>%
    select('month','Date','PM2.5')%>%
  group_by(month)%>%
  summarise(count=max(PM2.5,na.rm=TRUE))

p1<-ggplot(max_each_month,aes(as.numeric(month),count))+geom_line(color="magenta3")+
    scale_x_continuous(breaks=seq(1,12,1), limits=c(1,12)) +
    xlab('')+ylab('PM2.5') + ggtitle("Max for Each Month")

mean_each_month<-og_wanliu%>%
    select('month','Date','PM2.5')%>%
  group_by(month)%>%
  summarise(count=mean(PM2.5,na.rm=TRUE))

p2<-ggplot(mean_each_month,aes(as.numeric(month),count))+geom_line(color="magenta3")+
    scale_x_continuous(breaks=seq(1,12,1), limits=c(1,12)) +
    xlab('')+ylab('PM2.5') + ggtitle("Mean for Each Month")

min_each_month<-og_wanliu%>%
    select('month','Date','PM2.5')%>%
  group_by(month)%>%
  summarise(count=min(PM2.5,na.rm=TRUE))

p3<-ggplot(min_each_month,aes(as.numeric(month),count))+geom_line(color="magenta3")+
    scale_x_continuous(breaks=seq(1,12,1), limits=c(1,12)) +
    xlab('')+ylab('PM2.5') + ggtitle("Min for Each Month")

grid.arrange(p1,p2,p3)
```

- there is some kind of trend as the months progress. We see the numbers range is so different though

# Time Series on Bejing

```{r}
library(ggfortify)
PM25<-wanliu%>%
     as.data.frame %>%
      mutate(date_ymd=as.Date(Date,format="%Y-%m-%d"))%>%
           group_by(date_ymd)%>%
         summarise(max_pol=max(PM2.5,na.rm=TRUE))%>%
    as.data.frame %>%
    mutate(y_m=paste(substr(date_ymd,1,nchar(date_ymd)+2),"-01"))%>%
    group_by(y_m)%>%
   summarise(avg_emission=mean(max_pol,na.rm=TRUE))

PM25_flow<-data.frame("date"=as.Date(PM25$y_m,format="%Y - %m - %d"),"count"=pollutant$avg_emission)
  
#converting to time series
# count is avg emission
PM25_ts<-ts(PM25_flow$count,frequency = 12,start = c(2013,03))

#ggplot format of ts plot    
autoplot(PM25_ts,ts.colour ="black") +xlab('Date')+ylab('Count')+ggtitle('Time Series of PM2.5 Pollutant')
```

```{r}
PM25_ts_decompose<-decompose(PM25_ts)

actual<-autoplot(PM25_ts_decompose$x)+xlab("Year")+ylab("Count")+ggtitle("Actual time series of PM2.5")
seas<-autoplot(PM25_ts_decompose$seasonal)+xlab("Year")+ylab("Count")+ggtitle("Seasonality time series of PM2.5")
tren<-autoplot(PM25_ts_decompose$trend)+xlab("Year")+ylab("Count")+ggtitle("Trend time series of PM2.5")
grid.arrange(actual,seas,tren,ncol=1,top="Decomposition of PM2.5 Time Series")
```

- There is little seasonality effect and trend factor.




```{r}
plot(PM25_ts) # plot the raw data
abline(reg=lm(PM25_ts~time(PM25_ts))) # fit a trend line
```

```{r}
cycle(PM25_ts)
```


```{r}
boxplot(PM25_ts~cycle(PM25_ts))
```




```{r}

wanliu_ts <- as.ts(wanliu$PM2.5, frequency(24))
wanliu_ts %>% mstl() %>%
  autoplot() + xlab("Week")
```


```{r}
wanliu_ts_decompose<-decompose(wanliu_ts)

actual<-autoplot(wanliu_ts_decompose$x)+xlab("Year")+ylab("Count")+ggtitle("Actual time series of PM2.5")
seas<-autoplot(wanliu_ts_decompose$seasonal)+xlab("Year")+ylab("Count")+ggtitle("Seasonality time series of PM2.5")
tren<-autoplot(wanliu_ts_decompose$trend)+xlab("Year")+ylab("Count")+ggtitle("Trend time series of PM2.5")
grid.arrange(actual,seas,tren,ncol=1,top="Decomposition of PM2.5 Time Series")
```

```{r}
write.csv(wanliu,"C:\\Users\\isabellachittumuri\\Desktop\\Project\\STAT 715\\Wanliu.csv", row.names = FALSE)
```

```{r}
# save df into desktop
write.csv(wanliu,'wanliu.csv')
```

























