---
title: "715_project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Packages
library(dplyr)
library(tidyverse)
library(readxl)
library(readr)
library(tidyverse)
```

# Exploratory Analysis

Wanliu Cite 11 (PRSA_Data_Wanliu_20130301-20170228.csv) has the least amount of missing data with only 382 N.A. values

```{r}
wanliu <- read.csv("PRSA_Data_Wanliu_20130301-20170228.csv", header = T)
og_wanliu <- wanliu[c(2:6)]
```

```{r}
# combine year, month, day, hour
# in UTC (Coordinated Universal Time)
library(lubridate)
og_wanliu$Date <- with(og_wanliu, ymd_h(paste(year, month, day, hour, sep= ' '))) 
wanliu <- og_wanliu[c(5,6)] # only take pm2.5 and new date column
```

```{r}
# Precentage of NA values
nrow(wanliu) # 35064
summary(wanliu) # 382 NAs
# 382/35064 = 0.01089436 (about 1 precent)
```

The NAs only make up 1% of the data.

```{r}
# wanliu <- na.omit(wanliu) # omit missing values 
summary(wanliu)
str(wanliu)
class(wanliu)
```

Date column is in POSIXct format. 
We can see that there is a huge difference between median and maximum values.

# Check for outliers

## Before Aggregation

```{r}
library(gridExtra)

p1 <- wanliu %>%
  ggplot(aes(x = PM2.5)) +
  geom_boxplot(color = "brown", fill = "white") + labs(title="Distribution of PM2.5",x='ug/m3',y='PM2.5') 

p2 <- wanliu %>%
  ggplot(aes(x = PM2.5)) +
  geom_histogram(color = "brown", fill = "white", binwidth=20)

grid.arrange(p1,p2 ,nrow=2)
```


```{r}
mean_pollutants_per_day<-wanliu %>%
     as.data.frame %>%
    select('Date','PM2.5')%>%
      mutate(date_ymd=as.Date(Date,format="%Y-%m-%d"))%>%
           group_by(date_ymd)%>%
         summarise(PM2.5=mean(PM2.5,na.rm=TRUE))

daily_avg <- as.data.frame(mean_pollutants_per_day)
```

```{r}
p1 <- daily_avg %>%
  ggplot(aes(x = PM2.5)) +
  geom_boxplot(color = "goldenrod3", fill = "white") + labs(title="Distribution of PM2.5 on Daily Max Values",x='ug/m3',y='PM2.5') 

p2 <- daily_avg %>%
  ggplot(aes(x = PM2.5)) +
  geom_histogram(color = "goldenrod3", fill = "white", binwidth=20) + labs(x='ug/m3',y='Count')

grid.arrange(p1,p2 ,nrow=2)
```

# Time series flow

- From previous section, we saw the distribution of four pollutants. Now we will see the time series flow for the same.

```{r}
library(scales)
# using daily avg
ggplot(data=daily_avg,aes(x=date_ymd,y=PM2.5))+geom_line(color="lightslateblue")+
scale_x_date(breaks = seq(as.Date("2013-01-01"), as.Date("2017-01-01"), by="6 months"),labels = date_format("%b-%y"))+
xlab('Date')+ylab('ug/m3')+ggtitle(paste('Daily Maximum Emission of PM2.5'))+
theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")
```

# Time Series on Bejing

```{r}
library(ggfortify)
  
#converting to time series
# count is avg emission
PM25_ts<-ts(daily_avg$PM2.5,frequency = 365, start = c(2013,03))

PM25_ts <- as.data.frame(PM25_ts)

#ggplot format of ts plot    
autoplot(PM25_ts,ts.colour ="black") +xlab('Date')+ylab('Count')+ggtitle('Time Series of PM2.5 Pollutant')
```

```{r}
PM25_ts_decompose<-decompose(PM25_ts)

actual<-autoplot(PM25_ts_decompose$x)+xlab("Year")+ylab("Count")+ggtitle("Observed time series of PM2.5")
seas<-autoplot(PM25_ts_decompose$seasonal)+xlab("Year")+ylab("Count")+ggtitle("Seasonality time series of PM2.5")
tren<-autoplot(PM25_ts_decompose$trend)+xlab("Year")+ylab("Count")+ggtitle("Trend time series of PM2.5")
res<-autoplot(PM25_ts_decompose$random)+xlab("Year")+ylab("Count")+ggtitle("Residuals time series of PM2.5")
grid.arrange(actual,seas,tren,res,ncol=2,top="Decomposition of PM2.5 Time Series")
```

- There is seasonality and trend factor.

# Quantile Regression

```{r}
library(ggplot2)
ggplot(data=daily_avg,aes(x=date_ymd,y=PM2.5))+geom_line(color="black")+
scale_x_date(breaks = seq(as.Date("2013-01-01"), as.Date("2017-01-01"), by="6 months"),labels = date_format("%b-%y"))+
xlab('Date')+ylab('ug/m3')+ggtitle(paste('Daily Maximum Emission of PM2.5'))+
theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none") +
  geom_quantile(quantiles = 0.98)
```


```{r}
library(quantreg)
qr1 <- rq(PM2.5 ~ date_ymd, data=daily_avg, tau = c(0.25,0.5,0.75,0.9,0.95,0.98))
summary(qr1)

# Duplicate dataset, modify residuals
qr2 <- daily_avg %>%
  mutate(residuals=residuals(qr1), linpred=predict(qr1))

# Plot of predictions of model against residuals of model
qr2 %>%
  ggplot(aes(x = linpred, y = residuals)) +
  geom_point() +
  geom_smooth(method = 'loess')
```

```{r}
qs = c(0.25,0.5,0.75,0.9,0.95,0.98)
ggplot(data=daily_avg,aes(x=date_ymd,y=PM2.5))+geom_line(color="black")+
scale_x_date(breaks = seq(as.Date("2013-01-01"), as.Date("2017-01-01"), by="6 months"),labels = date_format("%b-%y"))+
xlab('Date')+ylab('ug/m3')+ggtitle(paste('24-Hour Emission of PM2.5'))+
theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none") +
  geom_quantile(quantiles = qs)
```





















